---
title: "Data Collection"
author: "Jonathan"
date: "2025-11-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Imports
```{r}
library(tidyverse)
library(readr)
library(dplyr)
setwd('/Users/jonat/Library/CloudStorage/OneDrive-ClarkUniversity/CU Courses/Economics Honors Thesis/Scripts')
```


# Data Collection

Here are the steps I need to do:

1.  DONE. ~~Go through the data dictionary and pick the variables I need to use. [Dictionary Link](https://nonprofit-open-data-collective.github.io/irs990efile/data-dictionary/data-dictionary.html)~~

2.  DONE. ~~Download the data .csvs for each part/section that I chose variables from. [Download Link](https://urbaninstitute.github.io/nccs/catalogs/catalog-efile-v2.html)~~

3.  DONE. ~~Download the csv for all the years.~~

4.  DONE. ~~I will need to go through each one and isolate the variables I want to keep,~~

5.  DONE. ~~Then save the keeped variables to a new data frame,~~

6.  DONE. ~~Then combine all the years, then save as a new/final .rds~~

7.  Start with some exploratory data analysis?

## Variables to Keep
My approach will be to find the folders of each category, then the files in each folder. From each of these files I need to extract only the variables of interest like below (below organized by category). Then combine these files so I have all the years of data for each category. Repeat this for each category. Then combine all the categories. So, by the end I will have one dataframe will all years data for all categories for all nonprofits of only the variables I want to keep. Then I will save this as a .csv for later.

```{r}
#1. I need to get the folders.
all_folders = list.dirs(path = "../Raw Data", recursive = FALSE)
print(all_folders)
```
Perfect.
```{r}
#2. I need to get the files, starting with just the first folder first as a test.
all_files_test = list.files(all_folders[1], recursive = FALSE, full.names = TRUE)
print(all_files_test)
```
Perfect.
```{r}
#3. Creating loop to extract all files from all folders.
all_files_list = lapply(all_folders, list.files, recursive = FALSE, full.names = TRUE)
print(all_files_list)
```
Perfect. Now I can iterate over this loop to complete all the tasks I need to do (extracting only variables of interest, saving to new csv)

## 1. Header
```{r}
list_of_dataframes_header = lapply(
  all_files_list[[1]], function(file_path) {
    
    #1. Adding year
    filename = basename(file_path)
    current_year = stringr::str_extract(file_path, "\\d{4}")
    
    #2. Selecting columns
    read_csv(file_path, show_col_types = FALSE) %>%
      select(
        ORG_EIN,
        F9_00_ORG_EIN,
        F9_00_ORG_ADDR_STATE,
        F9_00_ORG_ADDR_ZIP,
        F9_00_GRO_RCPT,
        # F9_00_EXEMPT_STAT_501C3_X, needs to be dropped
        F9_00_ORG_WEBSITE,
        F9_00_TYPE_ORG_CORP_X,
        F9_00_YEAR_FORMATION
      ) %>%
      mutate(Year = current_year) #adds new column
  }
)

# 2. Combine them into one data frame
header_data_combined = bind_rows(list_of_dataframes_header)

# 3. Check the result
print(tail(header_data_combined))
```
I will need to manually repeat these codes for each category, until I can combine them all. 

## 2. Summary
```{r}
# 1. Process all files in the *second* folder list
list_of_dataframes_summary = lapply(
  all_files_list[[2]], function(file_path) {
    
    # Extract the year
    current_year <- stringr::str_extract(file_path, "\\d{4}")
    
    # Read, select, and add the new 'Year' column
    read_csv(file_path, show_col_types = FALSE) %>%
      select(
        ORG_EIN,
        F9_01_ACT_GVRN_ACT_MISSION,
        F9_01_ACT_GVRN_TERMIN_KONTR_X,
        F9_01_ACT_GVRN_NUM_VOTE_MEMB,
        F9_01_ACT_GVRN_NUM_VOTE_MEMB_IND
      ) %>%
      mutate(Year = current_year) 
  }
)

# 2. Combine them into one data frame
summary_data_combined <- bind_rows(list_of_dataframes_summary)

# 3. Check the result
print(tail(summary_data_combined))
```
Perfect.
## 3. Mission
```{r}
# 1. Process all files in the *third* folder list
list_of_dataframes_mission = lapply(
  all_files_list[[3]], function(file_path) {
    
    # Extract the year
    current_year <- stringr::str_extract(file_path, "\\d{4}")
    
    # Read, select, and add the new 'Year' column
    read_csv(file_path, show_col_types = FALSE) %>%
      select(
        # F9_03_ORG_MISSION_PURPOSE, drop bc is redundant
        ORG_EIN,
        F9_03_PROG_NEW_X,
        F9_03_PROG_CHANGE_X
      ) %>%
      mutate(Year = current_year) 
  }
)

# 2. Combine them into one data frame
mission_data_combined <- bind_rows(list_of_dataframes_mission)

# 3. Check the result
print(tail(mission_data_combined))
```
Perfect.

## 4. Schedules (SKIPPED)
I am skipping this for now. It is not useful for the regression.

## 5. Governance
```{r}
# 1. Process all files in the *fifth* folder list
list_of_dataframes_governance <- lapply(
  all_files_list[[5]], function(file_path) {
    
    # Extract the first 4-digit number found in the file path
    current_year <- stringr::str_extract(file_path, "\\d{4}")
    
    # Read, select, and add the new 'Year' column
    read_csv(file_path, show_col_types = FALSE) %>%
      select(
        ORG_EIN,
        F9_06_GVRN_MEMB_STCKHLDR_X,
        F9_06_POLICY_JV_X
      ) %>%
      mutate(Year = current_year) 
  }
)

# 2. Combine them into one data frame
governance_data_combined <- bind_rows(list_of_dataframes_governance)

# 3. Check the result
print(tail(governance_data_combined))
```
Perfect.

## 6. Distribution to Trustees
```{r}
# 1. Process all files in the *sixth* folder list
list_of_dataframes_trustkey <- lapply(
  all_files_list[[6]], function(file_path) {
    
    # Extract the first 4-digit number found in the file path
    current_year <- stringr::str_extract(file_path, "\\d{4}")
    
    # Read, select, and add the new 'Year' column
    read_csv(file_path, show_col_types = FALSE) %>%
      select(
        ORG_EIN,
        F9_07_COMP_DTK_COMP_ORG_TOT,
        F9_07_COMP_DTK_COMP_RLTD_TOT,
        F9_07_COMP_DTK_COMP_OTH_TOT
      ) %>%
      mutate(Year = current_year) 
  }
)

# 2. Combine them into one data frame
trustkey_data_combined <- bind_rows(list_of_dataframes_trustkey)

# 3. Check the result
print(tail(trustkey_data_combined))
```
Perfect.

## 7. Revenue
```{r}
# 1. Process all files in the *seventh* folder list
list_of_dataframes_revenue <- lapply(
  all_files_list[[7]], function(file_path) {
    
    # Extract the first 4-digit number found in the file path
    current_year <- stringr::str_extract(file_path, "\\d{4}")
    
    # Read, select, and add the new 'Year' column
    read_csv(file_path, show_col_types = FALSE) %>%
      select(
        ORG_EIN,
        F9_08_REV_CONTR_FED_CAMP,
        F9_08_REV_CONTR_MEMBSHIP_DUE,
        F9_08_REV_CONTR_FUNDR_EVNT,
        F9_08_REV_CONTR_RLTD_ORG,
        F9_08_REV_CONTR_GOVT_GRANT,
        F9_08_REV_CONTR_OTH,
        F9_08_REV_CONTR_TOT,
        F9_08_REV_PROG_TOT_TOT,
        F9_08_REV_OTH_INVEST_INCOME_TOT,
        F9_08_REV_OTH_INVEST_BOND_TOT,
        F9_08_REV_OTH_ROY_TOT,
        F9_08_REV_OTH_RENT_NET_TOT,
        F9_08_REV_OTH_SALE_GAIN_NET_TOT,
        F9_08_REV_OTH_FUNDR_NET_TOT,
        F9_08_REV_OTH_GAMING_NET_TOT,
        F9_08_REV_OTH_INV_NET_TOT,
        F9_08_REV_MISC_TOT_TOT,
        F9_08_REV_TOT_TOT
      ) %>%
      mutate(Year = current_year) 
  }
)

# 2. Combine them into one data frame
revenue_data_combined <- bind_rows(list_of_dataframes_revenue)

# 3. Check the result
print(tail(revenue_data_combined))
```
Perfect.

## 8. Expenses
```{r}
# 1. Process all files in the *eighth* folder list
list_of_dataframes_expenses <- lapply(
  all_files_list[[8]], function(file_path) {
    
    # Extract the first 4-digit number found in the file path
    current_year <- stringr::str_extract(file_path, "\\d{4}")
    
    # Read, select, and add the new 'Year' column
    read_csv(file_path, show_col_types = FALSE) %>%
      select(
        ORG_EIN,
        F9_09_EXP_GRANT_US_ORG_TOT,
        F9_09_EXP_GRANT_US_INDIV_TOT,
        F9_09_EXP_GRANT_FRGN_TOT,
        F9_09_EXP_TOT_TOT,
        F9_09_EXP_TOT_PROG,
        F9_09_EXP_TOT_MGMT,
        F9_09_EXP_TOT_FUNDR
      ) %>%
      mutate(Year = current_year) 
  }
)

# 2. Combine them into one data frame
expenses_data_combined <- bind_rows(list_of_dataframes_expenses)

# 3. Check the result
print(head(expenses_data_combined))
```
Perfect.

# Combing Data
Now that I have 7 (not 8 anymore) data frames for each categories of variables, I want to combine them to one data frame with all my variables.

I realize I need to identify each non-profit organization with their EIN number. I will need to update my code to include that variable from each .csv.
```{r}
all_dataframe_list_list = list(
  header_data_combined,
  summary_data_combined,
  mission_data_combined,
  governance_data_combined,
  trustkey_data_combined,
  revenue_data_combined,
  expenses_data_combined
)

final_master_dataframe <- reduce(
  all_dataframe_list_list,
  full_join,
  by = c("ORG_EIN", "Year")
)

```

Double checking.
```{r}
final_master_dataframe = final_master_dataframe |>
  distinct(ORG_EIN, Year, .keep_all = TRUE)
```


# Saving Data
Perfect! Now saving.

```{r}
dir.create("Final_Data")
saveRDS(final_master_dataframe, "../Final_Data/final_master_data.rds")
```
