---
title: "1 Activity Type Data Collection"
author: "Jonathan"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('/Users/jonat/Library/CloudStorage/OneDrive-ClarkUniversity/CU Courses/Economics Honors Thesis/Scripts')
library(dplyr)
library(ggplot2)
library(tidyverse)
library(readr)
library(stringr)
options(scipen = 999)
```

## Import Data
First, I will need to combine my 3 data frames into one for the whole country.
```{r}
activity_data_1 = read_csv("../Raw Data/IRS Activity Codes/eo1.csv", 
                           col_types = cols(CLASSIFICATION = col_character()))

activity_data_2 = read_csv("../Raw Data/IRS Activity Codes/eo2.csv", 
                           col_types = cols(CLASSIFICATION = col_character()))

activity_data_3 = read_csv("../Raw Data/IRS Activity Codes/eo3.csv", 
                           col_types = cols(CLASSIFICATION = col_character()))

# Combining
activity_data = bind_rows(activity_data_1, activity_data_2, activity_data_3)

# Now, filter to just EIN and activity code variables ()
activity_data = activity_data %>%
  select(EIN, NTEE_CD, ACTIVITY) %>%
  mutate(EIN = as.numeric(EIN))
```

## Matching Categories
Right now, I have NTEE codes and activity codes, depending on the age of the non-profit organization. I want to translate them all to the same categories, using NTEE codes as the template since it is the modern version of activity categorizing for the IRS.
```{r}
library(dplyr)


# This function translates every activity_code (the old version) into new categories into a new column.
activity_data <- activity_data %>% # Thanks, AI. 
  mutate(
    activity_code = as.numeric(substr(ACTIVITY, 1, 3))
  ) %>%
  mutate(activity_category = case_when(
    !is.na(NTEE_CD) & NTEE_CD != "" ~ substr(NTEE_CD, 1, 1),
    # --- X: Religion-Related ---
    activity_code >= 1 & activity_code <= 29 ~ "X", #double-checked. 
    
    # --- B: Education ---
    activity_code >= 30 & activity_code <= 59 ~ "B", #double-checked. 
    
    # --- A: Arts, Culture & Humanities ---
    activity_code >= 60 & activity_code <= 149 ~ "A", #double-checked. 
    
    # --- E: Health Care (With Exceptions) ---
    activity_code == 166 ~ "F", #double-checked. 
    activity_code %in% c(161, 162) ~ "H", #double-checked. 
    activity_code >= 150 & activity_code <= 179 ~ "E", #double-checked. 
    
    # --- U: Science & Technology ---
    activity_code >= 180 & activity_code <= 199 ~ "U", #double-checked. 
    
    # --- S: Community Improvement ---
    activity_code >= 200 & activity_code <= 229 ~ "S", #double-checked. 
    
    # --- K: Food, Agriculture & Nutrition ---
    activity_code >= 230 & activity_code <= 249 ~ "K", #double-checked. 
    
    # --- Y: Mutual & Membership Benefit ---
    activity_code >= 250 & activity_code <= 279 ~ "Y", #double-checked. 
    
    # --- N: Recreation & Sports ---
    activity_code >= 280 & activity_code <= 319 ~ "N", #double-checked. 
    
    # --- O: Youth Development ---
    activity_code >= 320 & activity_code <= 349 ~ "O", #double-checked. 
    
    # --- C: Environment (With Exception) ---
    activity_code == 355 ~ "D", #double-checked. 
    activity_code >= 350 & activity_code <= 379 ~ "C", #double-checked. 
    
    # --- L: Housing & Shelter ---
    # Housing (380-399)
    activity_code >= 380 & activity_code <= 399 ~ "L", #double-checked. 
    
    # --- S: Community Improvement (Inner City) ---
    activity_code == 406 ~ "I", #double-checked. 
    activity_code == 407 ~ "M", #double-checked. 
    # Inner City/Community (400-429) -> S
    activity_code >= 400 & activity_code <= 429 ~ "S", #double-checked. 
    
    # --- R: Civil Rights & Advocacy ---
    # Civil Rights (430-459)
    activity_code >= 430 & activity_code <= 459 ~ "R", #double-checked. 
    
    # --- I: Crime & Legal-Related ---
    # Litigation (460-479)
    activity_code >= 460 & activity_code <= 479 ~ "I", #double-checked. 
    
    # --- R: Political/Legislative ---
    # Legislative (480-509)
    activity_code >= 480 & activity_code <= 559 ~ "R", #double-checked. 
    
    # --- P: Human Services (With Exception) ---
    # 573 is Rehab/Substance Abuse -> F (Mental Health)
    activity_code == 573 ~ "F", #double-checked. 
    activity_code == 572 ~ "I", #double-checked. 
    # Services to Individuals (560-579) -> P
    activity_code >= 560 & activity_code <= 575 ~ "P",  #double-checked. 
    
    # --- T: Philanthropy ---
    # Grantmaking (600-699)
    activity_code >= 600 & activity_code <= 699 ~ "T", #double-checked. 
    
    # --- Misc/Other (900+) ---
    activity_code == 900 ~ "Y", # Cemetery #double-checked. 
    activity_code == 907 | activity_code == 908 ~ "W", # Veterans/Patriotic -> Public Benefit #double-checked. 
    activity_code == 910 | activity_code == 911 ~ "Q", # Foreign #double-checked. 
    activity_code == 913 ~ "D", # Animal Cruelty #double-checked. 
    activity_code == 927 ~ "T", # Fundraising #double-checked. 
    
    # Default catch-all for any remaining codes
    TRUE ~ "Z"
  ))

```


## Saving Data
```{r}
# Now, filter to just EIN and activity code variables ()
activity_data = activity_data %>%
  select(EIN, activity_category) %>%
  mutate(EIN = as.numeric(EIN))
```

```{r}
saveRDS(activity_data, "../Final_Data/activity_data.rds")
```


