---
title: "1z Exploratory Analysis"
author: "Jonathan"
date: "`r Sys.Date()`"
output: html_document
---

# Imports
```{r setup, include=FALSE}
#Document setup
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)
```
```{r}
#Imports
library(dplyr)
library(ggplot2)
library(tidyverse)
library(readr)
library(stringr)
```
```{r}
#Data import
org_data = readRDS(
  "../Final_Data/clean_master_data.rds"
  )
```

# Data Exploration

I want to start with some fast facts of my data (length of years, number of unique organizations, number of organizations with data for all years).

In addition, some things I wish to explore of my key variables (program expense, government revenues, services revenues, and various fixed effects like age, state of HQ, and type of organization):

-   DONE. ~~number of observations per year~~
-   means, mins, max, and standard deviations
-   occurrences of null / NAs

Then, I will compile the results into a table and possibly a few graphics for my upcoming presentation.

## Meta Data Facts

```{r}
# number of observations, divided per year
unique_returns_by_year = org_data %>%
  # 1. Group the data by the Year column (using your specific column name)
  group_by(YEAR) %>%
  # 2. Summarize by counting the DISTINCT (unique) ORG_EIN values
  summarise(unique_returns = n_distinct(ORG_EIN))

plot_returns = ggplot(data= unique_returns_by_year, aes(x = factor(YEAR), y = unique_returns))+
  geom_col(fill = "purple") +
  geom_text(aes(label = unique_returns), vjust = -0.5) +
  labs(
    title = "Total number of Form 990 Returns",
    subtitle = "Summarized by year, from 2009-2023",
    y = "Number of returns",
    x = "Year"
  )
print(plot_returns)
```
Saving the plot.
```{r}
ggsave("../Outputs/total_returns.png", plot = plot_returns)

```

## Key Variables

### Defining
```{r}
key_variables = org_data %>% 
  select(
    # Demand side variables
    F9_09_EXP_GRANT_US_ORG_TOT,
    F9_09_EXP_GRANT_US_INDIV_TOT,
    F9_09_EXP_TOT_PROG,
    
    # Supply side variables
    F9_08_REV_CONTR_FED_CAMP, 
    F9_08_REV_CONTR_FUNDR_EVNT, 
    F9_08_REV_CONTR_GOVT_GRANT, 
    F9_08_REV_CONTR_MEMBSHIP_DUE, 
    F9_08_REV_PROG_TOT_TOT,
    
    #Fixed effect variables
    #F9_00_ORG_ADDR_STATE,
    #F9_00_YEAR_FORMATION,
    #F9_03_PROG_CODE
  )

# Temporary code fix to remove negatives in a column I forgot about.
key_variables = key_variables %>%
  filter(if_all(
    .cols = all_of("F9_08_REV_PROG_TOT_TOT"),
    .fns = ~ .x >= 0
  ))
```

### Summary Statistics
```{r}
library(psych)
summary_table = describe(key_variables)
print(summary_table)
```
```{r}
library(dplyr)

summary_table <- key_variables %>%
  summarise(across(
    everything(),
    list(
      N = ~sum(!is.na(.)),
      Mean = ~mean(., na.rm = TRUE),
      SD = ~sd(., na.rm = TRUE),
      Min = ~min(., na.rm = TRUE),
      Max = ~max(., na.rm = TRUE)
    )
  ))
```

```{r}
library(dplyr)
library(tidyr)

summary_table_tidy <- summary_table %>%
  pivot_longer(
    cols = everything(),
    names_to = c("variable", "stat"),
    names_sep = "_(?=[^_]+$)"
  ) %>%
  pivot_wider(
    names_from = stat,
    values_from = value
  )

```

```{r}
summary_table_tidy <- summary_table_tidy %>%
  mutate(variable = gsub("^F9_", "", variable))
```

```{r}
library(modelsummary)


datasummary_df(
  summary_table_tidy,
  output = "../Outputs/summary_statistics.docx"
)

```


# Exploring Data Filtering
## Program Expenses
Before I filter out any organizations based on not having program expenses, I want to check the distribution of this trend. 
```{r}
org_data <- org_data %>% arrange(ORG_EIN, YEAR)

# Step 2: Calculate the distribution of MAX consecutive zero-expense years
consecutive_dist <- org_data %>%
  filter(YEAR <= 2020) %>% # I want to keep the Covid effect
  group_by(ORG_EIN) %>%
  summarize(max_consecutive_zeros = {
    runs <- rle(F9_09_EXP_TOT_PROG == 0)
    # Identify the maximum length where the value is TRUE (i.e., it was a zero)
    if(any(runs$values)) max(runs$lengths[runs$values]) else 0
  }) %>%
  count(max_consecutive_zeros)

print(consecutive_dist)
```
From this table, I can see that there are a few thousand organizations that do not do any program expenses. Those are useless to my analysis and are arguable organizations that should not be in this filing category nor receiving tax exempt status. 
```{r}
#Visualization
org_data %>%
  group_by(ORG_EIN) %>%
  summarize(max_streak = {
    runs <- rle(F9_09_EXP_TOT_PROG == 0)
    if(any(runs$values)) max(runs$lengths[runs$values]) else 0
  }) %>%
  ggplot(aes(x = factor(max_streak))) +
  geom_bar(fill = "brown") +
  labs(title = "Distribution of Consecutive Zero-Expense Years",
       x = "Longest Consecutive Streak of $0 Expenses",
       y = "Number of Organizations") +
  theme_minimal()
```

## Revenue
```{r}
#Data summary of revenue
summary(org_data$F9_08_REV_TOT_TOT)
```
```{r}
#Data summary of log(revenue)
revenue = org_data$F9_08_REV_TOT_TOT
log_revenue = log(revenue + 1)
summary(log_revenue)
```
```{r}
#Visualization of revenue and log(revenue). Thanks, AI.
library(patchwork) # To put plots side-by-side

# Plot A: Raw Revenue (likely heavily skewed)
p1 <- ggplot(org_data, aes(x = revenue)) +
  geom_histogram(fill = "steelblue", bins = 50) +
  labs(title = "Raw Revenue", x = "Dollars ($)", y = "Count") +
  theme_minimal()

# Plot B: Log Revenue (likely more bell-shaped/normal)
p2 <- ggplot(org_data, aes(x = log_revenue)) +
  geom_histogram(fill = "darkorange", bins = 50) +
  labs(title = "Log-Transformed Revenue", x = "Log Value", y = "Count") +
  theme_minimal()

# Combine them
p1 + p2
```

```{r}
org_data <- org_data %>% arrange(ORG_EIN, YEAR)

# Step 2: Calculate the distribution of MAX consecutive zero-expense years
consecutive_dist_rev <- org_data %>%
  filter(YEAR <= 2020) %>% # I want to keep the Covid effect
  group_by(ORG_EIN) %>%
  summarize(max_consecutive_rev_zeros = {
    runs <- rle(F9_08_REV_TOT_TOT == 0)
    # Identify the maximum length where the value is TRUE (i.e., it was a zero)
    if(any(runs$values)) max(runs$lengths[runs$values]) else 0
  }) %>%
  count(max_consecutive_rev_zeros)

print(consecutive_dist_rev)
```
```{r}
#Visualization
org_data %>%
  group_by(ORG_EIN) %>%
  summarize(max_streak = {
    runs <- rle(F9_08_REV_TOT_TOT == 0)
    if(any(runs$values)) max(runs$lengths[runs$values]) else 0
  }) %>%
  ggplot(aes(x = factor(max_streak))) +
  geom_bar(fill = "darkgreen") +
  labs(title = "Distribution of Consecutive Zero-Revenue Years",
       x = "Longest Consecutive Streak of $0 Revenues",
       y = "Number of Organizations") +
  theme_minimal()
```


## Expenses and Revenue Outliers
```{r}
org_data <- org_data %>% arrange(ORG_EIN, YEAR)
# Step 2: Calculate the distribution of MAX consecutive zero-revenues years
consecutive_dist <- org_data %>%
  filter(YEAR <= 2020) %>% # I want to keep the Covid effect
  group_by(ORG_EIN) %>%
  summarize(max_consecutive_both_zeros = {
    both_zero = (F9_08_REV_TOT_TOT == 0 & F9_09_EXP_TOT_PROG == 0)
    runs <- rle(both_zero)
    # Identify the maximum length where the value is TRUE (i.e., it was a zero)
    if(any(runs$values)) max(runs$lengths[runs$values]) else 0
  }) %>%
  count(max_consecutive_both_zeros)

print(consecutive_dist)
```
```{r}
#Visualization
org_data %>%
  group_by(ORG_EIN) %>%
  summarize(max_streak = {
      both_zero = (F9_08_REV_TOT_TOT == 0 & F9_09_EXP_TOT_PROG == 0)
    runs <- rle(both_zero)
    if(any(runs$values)) max(runs$lengths[runs$values]) else 0
  }) %>%
  ggplot(aes(x = factor(max_streak))) +
  geom_bar(fill = "blue") +
  labs(title = "Distribution of Consecutive Zero-Expense & Zero-Revenue Years",
       x = "Longest Consecutive Streak of $0 Values",
       y = "Number of Organizations") +
  theme_minimal()
```

