---
title: "1z Exploratory Analysis"
author: "Jonathan"
date: "`r Sys.Date()`"
output: html_document
---

# Imports

Document set-up.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('/Users/jonat/Library/CloudStorage/OneDrive-ClarkUniversity/CU Courses/Economics Honors Thesis/Scripts')
options(scipen = 999)
```

Library imports.

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(readr)
library(stringr)
```

Raw Data import.

```{r}
org_data = readRDS(
  "../Final_Data/clean_master_data.rds"
  )
```

# Data Exploration

I want to start with some fast facts of my data (length of years, number of unique organizations, number of organizations with data for all years).

In addition, some things I wish to explore of my key variables (program expense, government revenues, services revenues, and various fixed effects like age, state of HQ, and type of organization):

-   DONE. ~~number of observations per year~~
-   means, mins, max, and standard deviations
-   occurrences of null / NAs

Then, I will compile the results into a table and possibly a few graphics for my upcoming presentation.

## Meta Data Facts

```{r}
# number of observations, divided per year
unique_returns_by_year = org_data %>%
  # 1. Group the data by the Year column (using your specific column name)
  group_by(YEAR) %>%
  # 2. Summarize by counting the DISTINCT (unique) ORG_EIN values
  summarise(unique_returns = n_distinct(ORG_EIN))

plot_returns = ggplot(data= unique_returns_by_year, aes(x = factor(YEAR), y = unique_returns))+
  geom_col(fill = "purple") +
  geom_text(aes(label = unique_returns), vjust = -0.5) +
  labs(
    title = "Total number of Form 990 Returns",
    subtitle = "Summarized by year, from 2009-2023",
    y = "Number of returns",
    x = "Year"
  )
print(plot_returns)
```
Saving the plot.
```{r}
ggsave("../Outputs/total_returns.png", plot = plot_returns)

```

## Key Variables

### Defining
```{r}
key_variables = org_data %>% 
  select(
    # Demand side variables
    F9_09_EXP_GRANT_US_ORG_TOT,
    F9_09_EXP_GRANT_US_INDIV_TOT,
    F9_09_EXP_TOT_PROG,
    
    # Supply side variables
    F9_08_REV_CONTR_FED_CAMP, 
    F9_08_REV_CONTR_FUNDR_EVNT, 
    F9_08_REV_CONTR_GOVT_GRANT, 
    F9_08_REV_CONTR_MEMBSHIP_DUE, 
    F9_08_REV_PROG_TOT_TOT,
    
    #Fixed effect variables
    #F9_00_ORG_ADDR_STATE,
    #F9_00_YEAR_FORMATION,
    #F9_03_PROG_CODE
  )

# Temporary code fix to remove negatives in a column I forgot about.
key_variables = key_variables %>%
  filter(if_all(
    .cols = all_of("F9_08_REV_PROG_TOT_TOT"),
    .fns = ~ .x >= 0
  ))
```

### Summary Statistics
```{r}
library(psych)
summary_table = describe(key_variables)
print(summary_table)
```
```{r}
library(dplyr)

summary_table <- key_variables %>%
  summarise(across(
    everything(),
    list(
      N = ~sum(!is.na(.)),
      Mean = ~mean(., na.rm = TRUE),
      SD = ~sd(., na.rm = TRUE),
      Min = ~min(., na.rm = TRUE),
      Max = ~max(., na.rm = TRUE)
    )
  ))
```

```{r}
library(dplyr)
library(tidyr)

summary_table_tidy <- summary_table %>%
  pivot_longer(
    cols = everything(),
    names_to = c("variable", "stat"),
    names_sep = "_(?=[^_]+$)"
  ) %>%
  pivot_wider(
    names_from = stat,
    values_from = value
  )

```

```{r}
summary_table_tidy <- summary_table_tidy %>%
  mutate(variable = gsub("^F9_", "", variable))
```

```{r}
library(modelsummary)


datasummary_df(
  summary_table_tidy,
  output = "../Outputs/summary_statistics.docx"
)

```

