---
title: "3 Regression Analysis"
author: "Jonathan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('/Users/jonat/Library/CloudStorage/OneDrive-ClarkUniversity/CU Courses/Economics Honors Thesis/Scripts')
```

# Imports

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(readr)
library(stringr)
library(fixest)
library(stargazer)
```

```{r}
org_data = readRDS("../Final_Data/clean_master_data.rds")
```

# Model 1 (Demand Side Effect)

Y = Program Expenses (representing total demand for the non-profits services/goods)

X = GDP Growth (representing the effect of macro economic conditions, including recessions)

B1 = The effect of x amount of gdp growth on demand for the non-profit. *I would expect B1 to be negative. As the economy does better, the demand for services from non-profits (represented by Program Expenses, Y) decreases.*

U = Fixed effects of non-profit organizations.

```{r}
demand_side = org_data %>% 
  
  # Selecting the columns I need. Expense, identification, and year.
  select(F9_09_EXP_TOT_PROG, ORG_EIN, YEAR, real_gdp, gdp_change_percent, F9_08_REV_TOT_TOT ) %>%
  
  # Adding a new column for log(program_expenses) to make the data more interpretative
  mutate(Log_Expenses = log(F9_09_EXP_TOT_PROG + 1),
         Log_Revenues = log(F9_08_REV_TOT_TOT +1)) # +1 to eliminate errors
```

```{r}
# The regression model
demand_model = lm(Log_Expenses ~ gdp_change_percent, data = demand_side)

# Output
print(summary((demand_model)))
```

This regression states that when GDP does not change, so at the intercept when GDP change = 0%, Log(expenses) are equal to about 7.28.

For every 1% increase in GDP, Expenses decrease by about 5.4%. This is statistically significant. Though the very low R2 value implies this has little to no predictive power, but a correlation does exist.

```{r}
stargazer(demand_model, type = "text", summary = TRUE)
```

# Model 2 (Supply Side)

```{r}
# Selecting relevant variables
supply_side = org_data %>%
  select(ORG_EIN, YEAR, real_gdp, gdp_change_percent, F9_08_REV_CONTR_GOVT_GRANT,F9_08_REV_CONTR_FUNDR_EVNT,F9_08_REV_TOT_TOT) %>%
  
  # Creating log() values for variables, to improve statistical analysis
  mutate(Log_Gov_Revenues = log(F9_08_REV_CONTR_GOVT_GRANT +1),
         Log_Fundr_Revenues = log(F9_08_REV_CONTR_FUNDR_EVNT +1),
         Log_Revenues = log(F9_08_REV_TOT_TOT +1)
         )
```

```{r}
# The regression model
supply_model_gov = lm(Log_Gov_Revenues ~ gdp_change_percent, data = supply_side)

# Output
summary(supply_model_gov)
```

## Interpretation: Government

This regression model tells me that for every 1% increase in GDP, Log(Government Revenues) for non-profits declines by 5.9%. This result is statistically significant correlation, but with little to no predictive power.

```{r}
# The regression model
supply_model_fundr = lm(Log_Fundr_Revenues ~ gdp_change_percent, data = supply_side)

# Output
summary(supply_model_fundr)
```

## Interpretation: Fundraising

# Model 3 (Controlled)

In this model, I will add a variable to the regression for non-profit revenues, to isolate the effect of the macroeconomic conditions, or GDP change percent, from its indirect effect on non-profit organizations.

```{r}
# Selecting relevant variables
revenue_controlled = org_data %>%
  select(ORG_EIN, real_gdp, gdp_change_percent, YEAR, F9_08_REV_TOT_TOT, F9_09_EXP_TOT_PROG) %>%
  # creating log values for better analysis
  mutate(Log_Expenses = log(F9_09_EXP_TOT_PROG + 1),
         Log_Revenues = log(F9_08_REV_TOT_TOT +1)) # +1 to eliminate errors)
```

```{r}
# regression model
revenue_controlled_model = lm(Log_Expenses ~ gdp_change_percent + Log_Revenues, data = revenue_controlled)

# output
summary(revenue_controlled_model)
```

# Model 4 (Fixed effects)

```{r}
# Selecting relevant variables
revenue_controlled = org_data %>%
  select(ORG_EIN, real_gdp, gdp_change_percent, YEAR, F9_08_REV_TOT_TOT, F9_09_EXP_TOT_PROG, activity_category) %>%
  # creating log values for better analysis
  mutate(Log_Expenses = log(F9_09_EXP_TOT_PROG + 1),
         Log_Revenues = log(F9_08_REV_TOT_TOT +1)) # +1 to eliminate errors)
```

```{r}
# regression model
revenue_controlled_model = feols(Log_Expenses ~ gdp_change_percent * activity_category, data = revenue_controlled)

# output
summary(revenue_controlled_model)
```
