---
title: "3b Preliminary Results"
author: "Jonathan"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Imports

```{r}
data_set = readRDS("../Final_Data/final_data_set.rds")

#Packages
library(dplyr)
library(fixest)
```

## Regression 1

The first regression will be as simple as possible to find any sort of correlation.

Log(Program Expenses) on GDP Change %. With macro fixed effects (state, year)

```{r}
#Setup code
library(dplyr)

regression1_variables = data_set %>%
  select(total_program_expenses, gdp_change_percent, year, state
         ) %>%
  mutate(
    log_program_expenses = log(total_program_expenses + 1)
  )

#Regression code
model1 = feols(log_program_expenses ~ gdp_change_percent | year + state, data = regression1_variables)

#Output code
summary(model1)
```

This regression tells me that for every 1 unit change in real GDP (1% increase), is correlated with a 0.7% decrease in total program expenses. This result is statistically significant.

## Regression 2

Now, I want to add a variable for each industry of non-profits, to add industry fixed effects.

```{r}
#Setup code
library(dplyr)
regression2_variables = data_set %>%
  select(total_program_expenses, gdp_change_percent, year, state, 
         industry #new
         ) %>%
  mutate(
    log_program_expenses = log(total_program_expenses + 1)
  )
#Regression code
model2 = feols(log_program_expenses ~ gdp_change_percent | year + state + industry, data = regression2_variables)

#Output code
summary(model2)
```

This regression tells me that for every 1 unit (1% increase) in real GDP, total program expenses decrease by about 0.5%. This result is not statistically significant though.

## Regression 3

Now, I want to include organization fixed effects. Controlling for each individual non-profit and comparing to its own finances.

```{r}
#Setup code
library(dplyr)
regression3_variables = data_set %>%
  select(total_program_expenses, gdp_change_percent, year, state, industry,
         organization_ein #new
         ) %>%
  mutate(
    log_program_expenses = log(total_program_expenses + 1)
  )
#Regression code
model3 = feols(log_program_expenses ~ gdp_change_percent | year + state + industry + organization_ein, data = regression3_variables)

#Output code
summary(model3)
```

This regression tells me that when controlling for each organization, a 1 unit increase (1% increase) in real GDP, is correlated with a 0.6% increase in program expenses. This result is statistically significant.

In other words, there appears to a cyclical trend with GDP and program expenses. As the economy expands, so do non-profit activities.

## Regression 4

Next, I will hold total revenues of the organization as fixed, by controlling it as a variable in the regression, to eliminate any effect that changing budgets has on changing program expenses. By isolating this effect, I (presumably) am left with only how much changing program expenses depends on economic conditions, and not budget conditions anymore.

In other words, I am looking for B\<0, because this would convey that as economic conditions improve, demand for non-profits decrease and so do program expenses.

`Arguing this correlation also exists backwards (worsening conditions leads to increased demand and increased expenses) is a big assumption, but a claim I want to reach with my analysis.`

```{r}
#Setup code
library(dplyr)
regression4_variables = data_set %>%
  select(total_program_expenses, gdp_change_percent, year, state, industry,
         total_revenue #new
         ) %>%
  mutate(
    log_program_expenses = log(total_program_expenses + 1),
    log_revenue = log(total_revenue +1)
  )
#Regression code
model4 = feols(log_program_expenses ~ gdp_change_percent + log_revenue | year + state + industry, data = regression4_variables)

#Output code
summary(model4)
```

This regression tells me that for every 1 unit (1%) increase in real GDP, program expenses decrease by about 0.2%, though this result is not statistically significant. Additionally, for every 1 unit (1%) increase in revenues, program expenses increase by 0.84%, and this result is statistically significant.

In other words, for non-profit organizations of any sector or state, revenues has a stronger correlation in changing program expenses than economic conditions. Further supporting this is the correlation with GDP change, indicating that as the economy expands, revenues for organizations increase, allowing for increased program expansion / expenses.