---
title: "3a Preliminary Results"
author: "Jonathan"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Imports

```{r}
data <- readRDS("../Final_Data/final_data_set.rds")

# Packages
library(dplyr)
library(fixest)
library(modelsummary) # Best package for regression tables
library(flextable) # For Word tables
```

# Regressions

```{r}
regression_data <- data %>%
  select(total_program_expenses, total_revenue, gdp_change_percent, year, state, industry, organization_ein) %>%
  mutate(
    log_program_expenses = log(total_program_expenses + 1),
    log_revenues = log(total_revenue + 1),
    log_revenue = log(total_revenue + 1)
  )
```

## 1. Macro Fixed-effects

The first regression will be as simple as possible to find any sort of correlation. Log(Program Expenses) on GDP Change %. With macro fixed effects (state, year)

```{r}
# Regression code
macro_fixed_effects_demand_model <- feols(log_program_expenses ~ gdp_change_percent | year + state, data = regression_data)

macro_fixed_effects_supply_model <- feols(log_revenues ~ gdp_change_percent | year + state, data = regression_data)

# Output code
summary(macro_fixed_effects_demand_model) # Demand side
summary(macro_fixed_effects_supply_model) # Supply side
```

This regression tells me that for every 1 unit change in real GDP (1% increase), is correlated with a **0.7% decrease** in total program expenses. This result is statistically significant.

On the supply side (Revenue), a 1 unit change in real GDP (1% increase) is correlated with a **0.40% decrease** in total revenues ($p < 0.05$). This result is statistically significant.

## 2. Industry Fixed-effects

Now, I want to add a variable for each industry of non-profits, to add industry fixed effects.

```{r}
# Regression code
industry_fixed_effects_demand_model <- feols(log_program_expenses ~ gdp_change_percent | year + state + industry, data = regression_data)

industry_fixed_effects_supply_model <- feols(log_revenues ~ gdp_change_percent | year + state + industry, data = regression_data)

# Output code
summary(industry_fixed_effects_demand_model) # Demand
summary(industry_fixed_effects_supply_model) # Supply
```

This regression tells me that for every 1 unit (1% increase) in real GDP, total program expenses is correlated with a **0.5% decrease.** This result is **not statistically significant** though.

For revenues, a 1% increase in GDP is correlated with a **0.37% decrease**, but this result is **not statistically significant** ($p = 0.06$).

## 3. Organization Fixed-effects

Now, I want to include organization fixed effects. Controlling for each individual non-profit and comparing to its own finances.

```{r}
# Regression code
firm_fixed_effects_demand_model <- feols(log_program_expenses ~ gdp_change_percent | year + state + industry + organization_ein, data = regression_data)

firm_fixed_effects_supply_model <- feols(log_revenues ~ gdp_change_percent | year + state + industry + organization_ein, data = regression_data)

# Output code
summary(firm_fixed_effects_demand_model)
summary(firm_fixed_effects_supply_model)
```

This regression tells me that when controlling for each organization, a 1 unit increase (1% increase) in real GDP, is correlated with a **0.6% increase in program expenses.** This result is **statistically significant.**

Interestingly for revenues, when controlling for organization fixed effects, the relationship becomes positive: a 1% increase in GDP is correlated with a **0.64% increase** in total revenues ($p < 0.001$). This suggests that **within** an organization, revenues are pro-cyclical (move with the economy).

In other words, there appears to a cyclical trend with GDP and program expenses. As the economy expands, so do non-profit activities.

## 4. Revenue Controlled

Next, I will hold total revenues of the organization as fixed, by controlling it as a variable in the regression, to eliminate any effect that changing budgets has on changing program expenses. By isolating this effect, I (presumably) am left with only how much changing program expenses depends on economic conditions, and not budget conditions anymore.

In other words, I am looking for B\<0, because this would convey that as economic conditions improve, demand for non-profits decrease and so do program expenses.

`Arguing this correlation also exists backwards (worsening conditions leads to increased demand and increased expenses) is a big assumption, but a claim I want to reach with my analysis.`

```{r}
# Regression code
revenue_controlled_demand_model <- feols(log_program_expenses ~ gdp_change_percent + log_revenue | year + state + industry, data = regression_data)

# Output code
summary(revenue_controlled_demand_model)
```

This regression tells me that for every 1 unit (1%) increase in real GDP, program expenses **decrease by about 0.2%,** though this result is **not statistically significant.** Additionally, for every 1 unit (1%) increase in revenues, program expenses increase by 0.84%, and this result is statistically significant.

In other words, for non-profit organizations of any sector or state, revenues has a stronger correlation in changing program expenses than economic conditions. Further supporting this is the correlation with GDP change, indicating that as the economy expands, revenues for organizations increase, allowing for increased program expansion / expenses.

# Exporting

```{r}
model_list <- list(
  "Demand Side (Macro)" = macro_fixed_effects_demand_model,
  "Supply Side (Macro)" = macro_fixed_effects_supply_model,
  "Demand Side (Industry)" = industry_fixed_effects_demand_model,
  "Supply Side (Industry)" = industry_fixed_effects_supply_model,
  "Demand Side (Firm)" = firm_fixed_effects_demand_model,
  "Supply Side (Firm)" = firm_fixed_effects_supply_model,
  "Revenue Controlled" = revenue_controlled_demand_model
)

comparison_table <- modelsummary(
  model_list,
  stars = TRUE,
  gof_map = c("nobs", "r.squared", "FE: year", "FE: state", "FE: industry", "FE: organization_ein"),
  coef_rename = c(
    "gdp_change_percent" = "GDP Change %",
    "log_revenue" = "Log(Revenue)"
  ),
  title = "Non-Profit Economic Cycle: Demand-side vs. Supply-side",
  output = "flextable",
  font = "Times New Roman"
) %>%
  autofit() %>%
  theme_vanilla() %>%
  bold(part = "header")

save_as_docx(comparison_table, path = "../Outputs/preliminary_regression_table.docx")
```
```{r}
print(comparison_table)
```

