---
title: "3b Preliminary Results"
author: "Jonathan"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Imports
```{r}
org_data = readRDS("../Final_Data/clean_master_data.rds")

#Packages
library(dplyr)
```

# Regression 0.1
The first regression will be as simple as possible to find any sort of correlation. 

Log(Program Expenses) on GDP Change %.

```{r}
#Setup code
regression_variables = org_data %>%
  select(F9_09_EXP_TOT_PROG, gdp_change_percent) %>%
  
  rename( #renaming to match all conventions and make them easier to understand
    total_program_expenses = F9_09_EXP_TOT_PROG
  ) %>%
  
  mutate(
    log_program_expenses = log(total_program_expenses + 1)
  )
```

```{r}
#Regression code
model1 = lm(log_program_expenses ~ gdp_change_percent, data = regression_variables)
```

```{r}
#Output code
summary(model1)
```

This regression tells me that for every 1 unit change in real GDP (1% increase), is correlated with a 5.3% decrease in total program expenses. This result is statistically significant.  

# Regression 0.2
Now, I want to add a variable for each year of data, to control time-variant factors.
```{r}
#Setup code
regression_variables = org_data %>%
  select(F9_09_EXP_TOT_PROG, gdp_change_percent, YEAR) %>%
  
  rename( #renaming to match all conventions and make them easier to understand
    total_program_expenses = F9_09_EXP_TOT_PROG,
    year = YEAR
  ) %>%
  
  mutate(
    log_program_expenses = log(total_program_expenses + 1)
  )
```

```{r}
#Regression code
model2 = lm(log_program_expenses ~ gdp_change_percent + factor(year), data = regression_variables)
```

```{r}
#Output code
summary(model2)
```
This regression tells me that for every 1 unit (1% increase) in real GDP, total program expenses decrease by about 4.2%. This result is statistically significant. 

# Regression 0.3
Now, I want to include state-fixed effects to improve my model.
```{r}
#Setup code
regression_variables = org_data %>%
  select(F9_09_EXP_TOT_PROG, gdp_change_percent, YEAR, F9_00_ORG_ADDR_STATE) %>%
  
  rename( #renaming to match all conventions and make them easier to understand
    total_program_expenses = F9_09_EXP_TOT_PROG,
    year = YEAR,
    state = F9_00_ORG_ADDR_STATE
  ) %>%
  
  mutate(
    log_program_expenses = log(total_program_expenses + 1)
  )
```

```{r}
#Regression code
model3 = lm(log_program_expenses ~ gdp_change_percent + factor(year) + factor(state), data = regression_variables)
```

```{r}
#Output code
summary(model3)
```
This regression tells me that when controlling for states, the effect of state GDP is actually non existent, statistically. However, being in a different states does have a statistical effect on program expenses. Also, being in a different year does have a statistical effect on program expenses. 

# Regression 0.4
Now, I want to add industry fixed effects to my regression by controlling for industries of non-profits.
```{r}
#Setup code
regression_variables = org_data %>%
  select(F9_09_EXP_TOT_PROG, gdp_change_percent, YEAR, F9_00_ORG_ADDR_STATE, activity_category) %>%
  
  rename( #renaming to match all conventions and make them easier to understand
    total_program_expenses = F9_09_EXP_TOT_PROG,
    year = YEAR,
    state = F9_00_ORG_ADDR_STATE,
    industry = activity_category
  ) %>%
  
  mutate(
    log_program_expenses = log(total_program_expenses + 1),
    industry = toupper(trimws(industry))
    )
```

```{r}
#Regression code
model4 = lm(log_program_expenses ~ gdp_change_percent + factor(year) + factor(state) + factor(industry), data = regression_variables)
```

```{r}
#Output code
summary(model4)
```
This regression tells me that there is now a negative correlation again with state gdp, but it is not statistically significant. Also, being in certain industries has a statistically significant effect on program expenses. 

In other words, 

# Regression 1.0
Finally, I will hold total revenues of the organization as fixed, by controlling it as a variable in the regression, to eliminate any effeect that changing budgets has on changing program expenses. By isolating this effect, I (presumably) am left with out how much changing program expenses depends on responding to changing demand, since GDP raises demand for non-profits. 

```{r}
#Setup code
regression_variables = org_data %>%
  select(F9_09_EXP_TOT_PROG, gdp_change_percent, YEAR, F9_00_ORG_ADDR_STATE, activity_category, F9_08_REV_TOT_TOT) %>%
  
  rename( #renaming to match all conventions and make them easier to understand
    total_program_expenses = F9_09_EXP_TOT_PROG,
    year = YEAR,
    state = F9_00_ORG_ADDR_STATE,
    industry = activity_category,
    total_revenue = F9_08_REV_TOT_TOT
  ) %>%
  
  mutate(
    log_program_expenses = log(total_program_expenses + 1),
    industry = toupper(trimws(industry)),
    log_revenue = log(total_revenue + 1)
    )
```

```{r}
#Regression code
model5 = lm(log_program_expenses ~ gdp_change_percent + + log_revenue + factor(year) + factor(state) + factor(industry), data = regression_variables)
```

```{r}
#Output code
summary(model5)
```
This regression now controls for all the variables I can...

#Regression 2.0
To further improve my analysis, I want to control for each non-profit organization itself. To do this, I use a new regression package to compare each observation to itself.

```{r}
#Setup code
regression_variables = org_data %>%
  select(F9_09_EXP_TOT_PROG, gdp_change_percent, YEAR, F9_00_ORG_ADDR_STATE, activity_category, F9_08_REV_TOT_TOT, ORG_EIN) %>%
  
  rename( #renaming to match all conventions and make them easier to understand
    total_program_expenses = F9_09_EXP_TOT_PROG,
    year = YEAR,
    state = F9_00_ORG_ADDR_STATE,
    industry = activity_category,
    total_revenue = F9_08_REV_TOT_TOT,
    organization_ein = ORG_EIN
  ) %>%
  
  mutate(
    log_program_expenses = log(total_program_expenses + 1),
    industry = toupper(trimws(industry)),
    log_revenue = log(total_revenue + 1)
    )
```

```{r}
#Regression code
model6 = feols(log_program_expenses ~ gdp_change_percent + log_revenue | organization_ein + year + state + industry, data = regression_variables)
```
```{r}
#Output code
summary(model6, cluster = ~organization_ein)
```

