---
title: "3e Covid Interaction Regressions"
author: "Jonathan"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Imports

```{r, message=FALSE, warning=FALSE}
# Load libraries
library(tidyverse)
library(fixest)
library(modelsummary)
library(flextable)

# Load data
# Pro-tip: use getwd() to verify your working directory
data_set <- readRDS("Final_Data/final_data_set.rds")
```

# Data Preparation

We need to create specific interaction variables:
1. **Post-Covid Indicator**: A dummy variable for years 2020 and later.
2. **Expanded During Covid**: An indicator for organizations that increased their revenue in 2020 compared to 2019.

```{r}
# 1. Identify organizations that expanded during Covid (2019 vs 2020)
org_expansion <- data_set %>%
  filter(year %in% c(2019, 2020)) %>%
  select(organization_ein, year, total_revenue) %>%
  # Handle duplicates if any (summing revenue for the year just in case)
  group_by(organization_ein, year) %>%
  summarise(total_revenue = sum(total_revenue, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = year, values_from = total_revenue, names_prefix = "rev_") %>%
  # Create the expansion dummy: 1 if 2020 revenue > 2019 revenue
  mutate(expanded_during_covid = ifelse(rev_2020 > rev_2019, 1, 0)) %>%
  select(organization_ein, expanded_during_covid) %>%
  # Remove NAs (orgs that didn't exist in both years)
  filter(!is.na(expanded_during_covid))

# 2. Merge back and create variables for regression
regression_data <- data_set %>%
  # Join the expansion info (left join keeps all orgs, but non-matching ones get NA for expansion)
  left_join(org_expansion, by = "organization_ein") %>%
  mutate(
    # Log transformations
    log_program_expenses = log(total_program_expenses + 1),
    log_revenue = log(total_revenue + 1),

    # Post-Covid Dummy (Year >= 2020)
    post_covid = ifelse(year >= 2020, 1, 0)
  )

# Check how many organizations we successfully labeled
cat("Number of organizations with expansion data:", sum(!is.na(regression_data$expanded_during_covid)), "\n")
```

# Regressions

The base model requested is:
`feols(log_program_expenses ~ gdp_change_percent + log_revenue | year + state + industry)`

We will add interactions to this base.

## Model 1: Interaction with Covid Era (Before/After)

We interact `gdp_change_percent` with `post_covid` to see if the relationship changed after the pandemic started.

*Note: The main effect of 'post_covid' is absorbed by the 'year' fixed effects, so it will be dropped, but the interaction will remain.*

```{r}
model_covid_interaction <- feols(
  log_program_expenses ~ gdp_change_percent * post_covid + log_revenue | year + state + industry,
  data = regression_data
)

summary(model_covid_interaction)
```

## Model 2: Interaction with Expanding Organizations

We interact `gdp_change_percent` with `expanded_during_covid` to see if organizations that grew during the pandemic respond differently to GDP changes.

```{r}
model_expansion_interaction <- feols(
  log_program_expenses ~ gdp_change_percent * expanded_during_covid + log_revenue | year + state + industry,
  data = regression_data
)

summary(model_expansion_interaction)
```

## Model 3: Both Interactions

Including both interaction terms in the same model.

```{r}
model_combined <- feols(
  log_program_expenses ~ gdp_change_percent * post_covid +
    gdp_change_percent * expanded_during_covid +
    log_revenue | year + state + industry,
  data = regression_data
)

summary(model_combined)
```

# Export Results

Creating a comparison table of the models.

```{r}
models_list <- list(
  "Covid Interaction" = model_covid_interaction,
  "Expansion Interaction" = model_expansion_interaction,
  "Combined" = model_combined
)

# Create the summary table
interaction_table <- modelsummary(
  models_list,
  stars = TRUE,
  gof_map = c("nobs", "r.squared", "adj.r.squared", "FE: year", "FE: state", "FE: industry"),
  coef_rename = c(
    "gdp_change_percent" = "GDP Change %",
    "log_revenue" = "Log(Revenue)",
    "post_covid" = "Post-Covid",
    "gdp_change_percent:post_covid" = "GDP % × Post-Covid",
    "expanded_during_covid" = "Expanded Org",
    "gdp_change_percent:expanded_during_covid" = "GDP % × Expanded Org"
  ),
  title = "Regression Results: Covid & Expansion Interactions",
  output = "flextable"
) %>%
  autofit() %>%
  theme_vanilla() %>%
  bold(part = "header")

# Save to Word
save_as_docx(interaction_table, path = "Outputs/covid_interaction_regressions.docx")
cat("✓ Table saved to Outputs/covid_interaction_regressions.docx\n")
```
